---
title: "Modelling of a length-based multi-species population targeted by a mixed fishery"
author: "Laurent Dubroca"
date: "`r Sys.Date()`"
output: html_document
bibliography: '/home/moi/Zotero/exportdb/My Library.bib'
---

# Framework

This report runs a length-based multi-species population model targeted by a
mixed fishery following @spence2020 (based on @hall2006). 
The model choice is related to the need of having a population structured in length
in order to then simulate sampling in it, and to have a model easily accessible in R.
The fishery catches are modelled by gear, and species preference and catchability
are highly customisable. These conditions are necessary to simulate realistic concurrent 
sampling plan.

# Package and co

Library are loaded and code sourced, if any.

```{r setup, include=T}
#link https://github.com/CefasRepRes/LeMaRns 
#https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0227767
library(LeMaRns)
#basic stuff
library(ggplot2)
library(dplyr)
```

# Population and fishery modelling

## Simulation test

To test the model, the example input parameters from @thorpe2015 are used (21 species, 50 years), with a constant fishing effort.
The SSB is plotted.

```{r runexample, include=T,eval=T}
#a short test
NS_params <- LeMansParam(NS_par, tau=NS_tau, eta=NS_eta, L50=NS_L50,other=1e12)
#NS_par: species growth and co
#tau: interaction matrix (row prey columns predator if 1 ij means j eat i)
#eta: steepness of the catchability curve
#L50: length @50% of the max catchability of the catchability curve
#other: amount of other food available for prey not in the model (in g)
#mortality: background mortality M1, pred mortality M2, fishing mort F
#see help for the description for the other parameters
#effort
effort <- matrix(0.25, 50, dim(NS_params @Qs)[3])
#4 gears
model_run <- run_LeMans(NS_params, years=50, effort=effort)
plot_SSB(NS_params,model_run)
```

## Simulation : biology and fishery

The biological model follows the example given by the library.
The fishery models follows the package's vignette example : 4 gears with different species
preference and catchability curves are used at a very low effort (0.25 at the model scale).
This low level of exploitation let the biological model fluctuates only based on the population processes.

```{r bunbioeff, include=T,eval=T}
## mixed fishery
## setup the model
NS_params <- LeMansParam(df=NS_par, gdf = NS_mixed_fish, tau=NS_tau, eta=NS_eta, L50=NS_L50, other=NS_other)
#gear and effort definition (0 -> no fishery)
effort <- matrix(0, 50, dim(NS_params@Qs)[3])
colnames(effort) <- c("Industrial", "Otter", "Beam", "Pelagic")
### run the initial 50 years with no fishery
model_run <- run_LeMans(NS_params, years=50, effort=effort)
#define a very low fishing effort
ef<-as.numeric(rep(0.25,4))
effort_mat <- matrix(ef, 50, dim(NS_params@Qs)[3], byrow=T)
colnames(effort_mat) <- c("Industrial", "Otter", "Beam", "Pelagic")
### run the model
model_run <- run_LeMans(params=NS_params, N0=model_run@N[,,501], years=50, effort=effort_mat)
#plot SSB and biomass
plot_SSB(inputs=NS_params,outputs=model_run,full_plot_only=T)
#png("rezbiomass.png")
plot_biomass(inputs=NS_params,outputs=model_run,full_plot_only=T)
#dev.off()

#get annual catch per gear dim i species j gear k timestep
ycatch<-data.frame((get_annual_catch(NS_params,outputs=model_run)))
colnames(ycatch)<-NS_params@species_names
ycatch$year<-1:50
pipo<-tidyr::pivot_longer(ycatch,cols=1:length(NS_params@species_names))
ggplot(pipo,aes(x=year,y=value,color=name))+geom_line()

#get catch per gear
cpg0<-get_CPG(inputs=NS_params,outputs=model_run,effort=effort_mat,years=50)
#add names to pivot_longer the array
attr(cpg0,"dimnames")[[1]]<-NS_params@species_names
attr(cpg0,"dimnames")[[2]]<-names(data.frame(effort_mat))
phi_min<-0.1
nbstep<-dim(cpg0)[3]
nbyear<-(nbstep)*phi_min
idyear<-ceiling(((1:(nbyear/phi_min))*phi_min))
attr(cpg0,"dimnames")[[3]]<-paste0("Y",idyear)
cpg0long<-data.frame(ftable(cpg0))%>%
	transmute(spp=as.character(Var1),
		  year=as.numeric(sub("Y","",as.character(Var3))),
		  gear=as.character(Var2),
		  value=Freq)%>%
	group_by(spp,year,gear)%>%summarise(value=sum(value,na.rm=T))%>%
	ungroup()
#png("rezcatch.png")
ggplot(cpg0long,aes(x=year,y=value,color=spp))+geom_line()+facet_wrap(~gear)
#dev.off()

#build a function to get number at len by gear {{{
#parameters
get_NPG<-function(Catch,Qs,effort,years=nrow(effort),phi_min=0.1){
#if(F){
#Catch<-model_run@N
#Qs<-NS_params@Qs
#effort<-effort_mat
#years<-nrow(effort)
#ph_min<-0.1
#}
	Catch<-Catch[,,-1]
        tot_time <- dim(Catch)[3]
        Fs <- array(0, dim=c(dim(Catch)[1], dim(Catch)[2], tot_time,dim(Qs)[3]))
        year <- ceiling(1:tot_time*phi_min)
	dimnames(Fs)[[4]] <- colnames(effort)
	for (j in colnames(effort)) {
	  for (ts in 1:tot_time) {
	    Fs[,,ts,j] <- effort[year[ts], j]*Qs[,,j]
	  }
	}
	#norm of fishing mort
 	tele0<-Fs
            for (ts in 1:tot_time) {
              tmp <- apply(Fs[,,ts,], c(1,2), sum)
              for(j in 1:dim(tele0)[4]) {
                tele <- Fs[,,ts,j]/tmp
                if (any(is.na(tele))) {
                  tele[is.na(tele)] <- 0
                }
		tele0[,,ts,j]<-tele
              }
	    }
	#get N fish by gear
	NPG<-tele0
        for(j in 1:dim(tele0)[4]) {
		NPG[,,,j]<-tele0[,,,j]*Catch
	}
	attr(NPG,"dimnames")[[1]]<-paste0("L",NS_params@u_bound)
	attr(NPG,"dimnames")[[2]]<-NS_params@species_names
	nbstep<-dim(NPG)[3]
	nbyear<-(nbstep)*phi_min
	idyear<-ceiling(((1:(nbyear/phi_min))*phi_min))
	attr(NPG,"dimnames")[[3]]<-paste0("Y",idyear)
	return(NPG)
}#}}}
npg0<-get_NPG(Catch=model_run@N,Qs=NS_params@Qs,effort=effort_mat,years=nrow(effort),phi_min=0.1)

npg0long<-data.frame(ftable(npg0))%>%
	transmute(len=as.numeric(sub("L","",as.character(Var1))),
		  spp=as.character(Var2),
		  year=as.numeric(sub("Y","",as.character(Var3))),
		  gear=as.character(Var4),
		  value=Freq)%>%
	group_by(len,spp,year,gear)%>%summarise(value=sum(value,na.rm=T))%>%
	ungroup()
#plot the len distrib by year/gear/years
ggplot(npg0long,aes(x=len,y=value,color=year,group=year))+geom_line()+facet_grid(spp~gear,scales="free")
#ggsave("reznlenpop.png")
```

## Simulation : detailed fishery activities 

As it is, LeMaRns models produces bulk catches by gear.
The following code details the fishery catches by vessels, trips and fishing operation.
These units are generated randomly following some inputs parameters. 
In this report there are
4 Beam, 2 Industrial, 4 Otter and 3 Pelagic vessels.
Each of them do randomly 1 to 20 trips a year, in which 1 to 20 fishing operations occurs.
The number of active vessels by year vary from 1 to the maximum of number of the considered gear.
Biological parameters are added to the simulation (fish weights, ages...).
The catch for each fishing operation is a fraction of the total catch of the
gear, taken randomly and normalized to keep the total catch of the gear equal to
the LeMaRns total catches.

```{r simusamp,eval=T}
#improvment: generating trrrrip for the whole pop and sampling theeeeeeem
#npg0long,NS_params,n1=20,nbvess=data.frame(gear=c("Beam","Industrial","Otter","Pelagic"),nvess=c(4,2,4,3))
#ntrip is 1:n1
	#add vessel vessel trip proportionnal
	tr0<-npg0long%>%select(year,gear)%>%distinct()
	#generate a vessel population and related trip
	nbvess<-data.frame(gear=c("Beam","Industrial","Otter","Pelagic"),nvess=c(4,2,4,3))
	#generate up to 20 FO by gear and gear and up to max fo trip
	nfo<- tr0%>%rowwise()%>%mutate(nfo=sample(1:20,1))%>%
		mutate(ntrip=sample(1:nfo,1,replace=T))%>%
		ungroup()%>%
		left_join(nbvess)
	#trip id + FO id
	nfo<-nfo[rep(1:nrow(nfo),times=nfo$nfo),]%>%
		group_by(year,gear,nfo)%>%
		mutate(FOid=row_number())%>%ungroup()%>%
		rowwise()%>%
		mutate(TRid=sample(1:ntrip,1,replace=TRUE))%>%
		group_by(year,gear,TRid)%>%
		mutate(FOid=row_number())%>%ungroup()%>%
		arrange(year,gear,TRid,FOid)%>%
		ungroup()
	#separate elaboration of vess id (unique by trip: trick for that)
	nvesstmp<-nfo%>%select(year,gear,TRid,nvess)%>%distinct()%>%
		rowwise()%>%mutate(VDid=sample(1:nvess,1,replace=T))%>%
		ungroup()
	nall<-nfo%>%left_join(nvesstmp)
	#generate the prop of the catch taken by each FO
	nall<-nall%>%group_by(year,gear)%>%mutate(prop=runif(nfo))%>%
		mutate(prop=prop/sum(prop))%>%ungroup()
	#pipo%>%group_by(year,gear)%>%summarise(tot=sum(prop))%>%filter(tot!=1)
	npg0longsim<-left_join(nall,npg0long)%>%mutate(value=value*prop)%>%
		filter(value>0)%>%select(-prop)
	#add nspp tot
	npg0longsim<-npg0longsim%>%
		group_by(year,gear,FOid,TRid,VDid)%>%
		mutate(nspp=n_distinct(spp))%>%
		group_by(year,gear,TRid,VDid)%>%
		mutate(nfo=n_distinct(FOid))%>%
		group_by(year,gear,VDid)%>%
		mutate(ntrip=n_distinct(TRid))%>%
		group_by(year,gear)%>%
		mutate(nvess=n_distinct(VDid))%>%
		ungroup()


	#check
	pipo1<-npg0longsim%>%group_by(year,gear,len,spp)%>%summarise(value0=sum(value))%>%ungroup()
	uu<-full_join(pipo1,npg0long)%>%filter(abs(value0-value)>1)
	if(nrow(uu)>0){strop("pb trip prop")}
	#npg0longsim%>%filter(year==1,gear=="Beam",FOid==1,TRid==1,VDid==1)
	#add biological information
	lwa<-data.frame(spp=NS_params@species_names,a=NS_params@W_a,b=NS_params@W_b,
			Linf=NS_params@Linf,k=NS_params@k)
	npg0longsim<-npg0longsim%>%left_join(lwa)
	npg0longsim$wind<-npg0longsim$a*npg0longsim$len^npg0longsim$b
	#add age
	fct1<-function(t0=0,K,Linf,L){ max(1, ((-log(1 - L/Linf))/K + t0))}
	npg0longsim<-npg0longsim%>%mutate(lentmp=ifelse(len>Linf,Linf-1e-1,len))%>%
		rowwise()%>%
		mutate(age=fct1(0,k,Linf,lentmp))%>%
		ungroup()
	#ggplot(npg0longsim,aes(x=len,y=age,group=spp,color=spp))+geom_point()
	#add to w by spp
	npg0longsim<-npg0longsim%>%mutate(wspp=value*wind)
	#clean the data
	npg0longsim<-npg0longsim%>%transmute(year,gear,nspp,nfo,ntrip,nvess,FOid,TRid,VDid,spp,len,age,
					  n=value,wind,wspp)
	reztmp<-npg0longsim%>%select(year,gear,VDid,TRid,FOid)%>%distinct()%>%
		group_by(year,gear,VDid,TRid)%>%summarise(nfo=n_distinct(FOid))%>%
		ungroup()
	ggplot(reztmp,aes(x=year,y=nfo,group=TRid,color=TRid))+
		geom_path()+facet_grid(gear~VDid)+
		ggtitle("FO by trip, vessel and gear")
```

## Simulation : sampling the simulated fishery activities 

On the simulated detailed fishery activities, a yearly random selection is
operated on a fraction of the catches related to a fishing operation.
The sampling is stratified by gear (1 vessel by gear each year) and trips (1 trip). 
The fishing operations follows are
sampled systematically (1 every 3 operations are sampled). 15 species over the
21 available are selected randomly and 10 % of the corresponding
catches are reported as sampled. This scenario mimics at-sea sampling.

```{r simusampplan,eval=T}
	#sampling scheme
	#nb vessel, nbtrip ,n FO and prop sampled
	nplanVD<-2;nplanTR<-1;nplanFO<-seq(1,max(npg0longsim$nfo),2);
	nplanSPP<-15;planprop<-0.1
	#sample space 
	sampspace<-npg0longsim%>%select(year,gear,spp,FOid,TRid,VDid)%>%
		distinct()
	#random sampling of 2 vessels by strata
	sampVD<-sampspace%>%select(year,gear,VDid)%>%
		distinct()%>%
		group_by(year,gear)%>%
		slice_sample(n=nplanVD,replace=F)%>%
		ungroup()%>%
		mutate(VDsamp=TRUE)
	sampspace<-left_join(sampspace,sampVD)%>%mutate(VDsamp=ifelse(is.na(VDsamp),F,VDsamp))
	#weighted random sampling of 1 trip by vessels by strata (aka the vessel
	#with the more trip are highest proba of selection)
	sampTR<-sampspace%>%filter(VDsamp==T)%>%select(year,gear,VDid,TRid)%>%
		distinct()%>%
		group_by(year,gear,VDid)%>%mutate(nbtrip=n_distinct(TRid))%>%
		group_by(year,gear,VDid)%>%
		slice_sample(n=nplanTR,weight_by=nbtrip,replace=F)%>%
		ungroup()%>%
		mutate(TRsamp=TRUE)%>%
		select(-nbtrip)
	sampspace<-left_join(sampspace,sampTR)%>%mutate(TRsamp=ifelse(is.na(TRsamp),F,TRsamp))
	#take one FO on 3
	sampspace<-sampspace%>%
		mutate(FOsamp=VDsamp&TRsamp&FOid%in%nplanFO)
	#plan to sample 10% of the catch
	sampspace<-sampspace%>%
		mutate(propsamp=ifelse(FOsamp,planprop,0))
	#take n species over the whole species there
	sampSPP<-sampspace%>%filter(VDsamp==T&TRsamp&FOsamp==T)%>%select(year,gear,VDid,TRid,FOid,spp)%>%
		distinct()%>%
		group_by(year,gear,VDid,TRid,FOid)%>%mutate(nbspp=n_distinct(spp))%>%
		group_by(year,gear,VDid,TRid,FOid)%>%
		slice_sample(n=nplanSPP,replace=F)%>%
		ungroup()%>%
		mutate(SPPsamp=TRUE)%>%
		select(-nbspp)
	#sampSPP%>%filter(year==1,gear=="Beam")
	#merge
	sampspace<-left_join(sampspace,sampSPP)%>%mutate(SPPsamp=ifelse(is.na(SPPsamp),F,SPPsamp))

	#add the sampling plan to npglon
	npg0longsim<-left_join(npg0longsim,sampspace)
	#compute subtotal for some parameters
	npg0longsim<-npg0longsim%>%
		group_by(year,gear,FOid,TRid,VDid)%>%
		mutate(wFO=sum(wspp))%>%
		group_by(year,gear,TRid,VDid)%>%
		mutate(wTR=sum(wspp))%>%
		group_by(year,gear,VDid)%>%
		mutate(wVD=sum(wspp))%>%
		ungroup()
	#check
	#npg0longsim%>%filter(year==1,gear=="Beam",VDid==2)%>%select(wVD)%>%unique()%>%pull
	#npg0longsim%>%filter(year==1,gear=="Beam",VDid==2)%>%select(TRid,wTR)%>%distinct()%>%pull(wTR)%>%sum()
	#npg0longsim%>%filter(year==1,gear=="Beam",VDid==2)%>%select(FOid,wFO)%>%distinct()%>%pull(wFO)%>%sum()
	#sum(npg0longsim$wspp)
	#npg0longsim%>%select(gear,year,VDid,wVD)%>%distinct()%>%pull(wVD)%>%sum()
	#npg0longsim%>%select(gear,year,VDid,TRid,wTR)%>%distinct()%>%pull(wTR)%>%sum()
	#npg0longsim%>%group_by(year,gear,nfo,ntrip,nvess,FOid,TRid,VDid)%>%summarise(nbfo=n_distinct(FOid))

	#npg0longsim%>%select(year,gear,VDid,TRid,FOid,nspp,nfo,ntrip,nvess)%>%arrange(year,gear,VDid,TRid,FOid)%>%distinct()
	tmp<-npg0longsim%>%filter(FOsamp==T)%>%
		group_by(year,gear,len,spp)%>%
		summarise(n=sum(n*propsamp))%>%
		ungroup()
	#plot the len distrib by year/gear/years
	ggplot(tmp,aes(x=len,y=n,color=year,group=year))+
		geom_line(alpha=0.7)+
		facet_grid(spp~gear,scales="free")+
		ggtitle("Sampled length by year")+
		theme_bw()

```

# Outputs

Simulated population data and sampling are saved.

```{r output,eval=T}
saveRDS(npg0longsim,file="./outputs/datasimu.rds")
#truc for CL
saveRDS(cpg0long,file="./outputs/datapop.rds")
saveRDS(npg0long,file="./outputs/datapopn.rds")

```

# R session information

```{r Rsession,cache=TRUE,eval=T,echo=T,cache=T,include=T}
sessioninfo::session_info()
```

# References
