---
title: "Conversion of simulated to to the RDBES format"
author: "Laurent Dubroca"
date: "`r Sys.Date()`"
output: html_document
bibliography: '/home/moi/Zotero/exportdb/My Library.bib'
---

# Framework

In this report the data simulated by the script `LD_01_simulation.Rmd` are
exported to the RDBES format [@rdbes2021]. The format in the R environment is
defined by the code developped by David Currie [@currie2020] during WKEST and
POP past workshops.

# Package and co

Library are loaded and code sourced, if any.


```{r setup, include=T}
library(dplyr)
library(ggplot2)
#dave fct
source("RDBES_Functions.R")
```

# Conversion

Following @currie2020, an empty RDBES R object is generated and populated with
(1) the population data and (2) the sampled data. The hierarchy H1 is used.

```{r convdata, include=T}
# This file shows how to generate test data for the RDBES

# IMPORTANT: Hack to stop write.csv changing numbers to scientific notation
options(scipen=500) # big number of digits

## STEP 1) LOAD OUR DATA

# Load the validation data
validationData <- getValidationData(downloadFromGitHub = FALSE, fileLocation = './tableDefs/BaseTypes.xsd')
# Load the reference data
allowedValues <- loadReferenceData(downloadFromICES = FALSE)

# Load the lists of tables required for each hierarchy: either refresh from ICES
# or just use a local copy
allRequiredTables <- getTablesInHierarchies(downloadFromGitHub = FALSE, fileLocation = './tableDefs/')
#allRequiredTables <- getTablesInHierarchies(downloadFromGitHub = TRUE, fileLocation = './tableDefs/')


## STEP 2: generate a hierarchy

  # Define some parameters for our test data
  myHierarchyToGenerate <- 'H1'
  print(myHierarchyToGenerate)
  myLowerHierarchyToGenerate <- 'A'
  myYear <- 1965 
  myCountry <- 'ZZ'
  # Generate some random data
  myTestData <- createTestData(HierarchyToGenerate = "H1", 
			       LowerHierarchyToGenerate = "A", 
			       RDBESvalidationdata = validationData, 
			       RDBEScodeLists = allowedValues, 
			       RequiredTables = allRequiredTables, 
			       NumberOfStrata = list(DE=1,VS=1), 
			       NumberSampled = list(VS=1,FT=1,FO=1,SL=21,SA=1,FM=1),
			       NumberTotal = list(VS=1,FO=1,SS=1,FM=1,BV=1), 
			       SelectionMethods = list())
  myNewTestData <- makeTestDataMoreRealistic(DataToUse = myTestData,CountryToUse="ZZ",YearToUse="1965",MetierList= NULL,SpeciesList= NULL,RDBEScodeLists=allowedValues)
  #load simudat and add it to the myTestData

  # The data we just geenerated is too random and won't pass validation or upload check - lets fix that now
  datrdb<- makeTestDataMoreRealistic(DataToUse = myTestData,CountryToUse=myCountry,YearToUse=myYear,MetierList= NULL,SpeciesList= NULL,RDBEScodeLists=allowedValues)
  #use simudat to populate datrdb for all the year ?
  datsim<-readRDS("../outputs/datasimu.rds")


#a function to prepare datsim computing primary key and sample tot for datsim
# first step{{{
 sim2prepsim<-function(datsim,selec){
	 if(F){
	 selection<-datsim$FOsamp&datsim$SPPsamp
	 }
datsim$selec<-selec
  #prep pk and parameters inside the datsim object
  tmp<-datsim%>%
	  #FM
	  group_by(year,gear,FOid,TRid,VDid,spp,len)%>%
	  mutate(newFMid=cur_group_id())%>%ungroup()%>%
  	  #SA
	  mutate(Xwspp=ifelse(selec,wspp,NA),
		 Xspp=ifelse(selec,spp,NA))%>%
	  group_by(year,gear,FOid,TRid,VDid,spp)%>%
	  mutate(newSAid=cur_group_id(),
		 SAwtot=sum(wspp),SAwsamp=sum(Xwspp,na.rm=T),
		 SAntot=n_distinct(spp),SAnsamp=n_distinct(Xspp,na.rm=T),
		 SAsel="Census",SAname=paste(year,gear,VDid,TRid,FOid,spp,sep="_"))%>%
	  ungroup()%>% select(-Xwspp,-Xspp)%>%
	  #SS
	  group_by(year,gear,FOid,TRid,VDid)%>%
	  mutate( Xspp=ifelse(selec,spp,NA))%>%
	  mutate(newSSid=cur_group_id(),
		 SSntot=n_distinct(spp,na.rm=T),SSnsamp=n_distinct(Xspp,na.rm=T),
		 #SSsel="Census",SSname=paste(year,gear,VDid,TRid,FOid,spp,sep="_"))%>%#ungroup()
		 SSsel="Census",SSname=paste0(year,gear))%>%
	  ungroup()%>% select(-Xspp)%>%
	  #FO
	  mutate(XFOid=ifelse(selec,FOid,NA))%>%
	  group_by(year,gear,FOid,TRid,VDid)%>%
	  mutate(newFOid=cur_group_id())%>%
	  group_by(year,gear,TRid,VDid)%>%
	  mutate(FOntot=n_distinct(FOid,na.rm=T),FOnsamp=n_distinct(XFOid,na.rm=T))%>%ungroup()%>%
	  mutate(FOsel="Census",FOname=paste(year,gear,VDid,TRid,FOid,sep="_"),
		 FOstratname=paste(year,gear,VDid,TRid,sep="_"))%>%
	  ungroup()%>%select(-XFOid)%>%
	  #FT
	  mutate(XTRid=ifelse(selec,TRid,NA))%>%
	  group_by(year,gear,TRid,VDid)%>%
	  mutate(newFTid=cur_group_id())%>%
	  group_by(year,gear,VDid)%>%
	  mutate(FTntot=n_distinct(TRid,na.rm=T),
		 FTnsamp=n_distinct(XTRid,na.rm=T))%>%ungroup()%>%
	  mutate(FTsel="Census",FTname=paste(year,gear,VDid,TRid,sep="_"),
		 FTstratname=paste(year,gear,VDid,sep="_"))%>%
	  ungroup()%>%select(-XTRid)%>%
	  #VS
	  mutate(XVDid=ifelse(selec,VDid,NA))%>%
	  group_by(year,gear,VDid)%>%
	  mutate(newVSid=cur_group_id())%>%
	  group_by(year,gear)%>%
	  mutate(VSntot=n_distinct(VDid,na.rm=T),VSnsamp=n_distinct(XVDid,na.rm=T))%>%ungroup()%>%
	  mutate(VSsel="Census",VSname=paste(year,gear,VDid,sep="_"),
		 VSstratname=paste(year,gear,sep="_"))%>%
	  ungroup()%>%select(-XVDid)%>%
	  #SD & DE
	  group_by(year)%>%
	  mutate(newSDid=cur_group_id(),newDEid=cur_group_id())%>%
	  ungroup()
	  #tmp%>%
	#	  mutate(pk=paste(year,gear))%>%
	#	  group_by(pk,newVSid,VSntot,VSnsamp)%>%
	#	  summarise(SPPsamp=any(SPPsamp),FOsamp=any(FOsamp))%>%
	#	  ungroup()%>%arrange(pk,newVSid)%>%View
  	return(tmp)
  }
#}}}

#convert prepared simdat to rdbes
prepsim2rdbes<-function(datsim,datrdb,selec){#{{{
	if(F){
		tmp<-datsim
		selec<-datsim$FOsamp&datsim$SPPsamp
	}
  tmp<-datsim[selec,]
  #FM
  pipo<-tmp%>%transmute(newFMid,len,n,newSAid)%>%distinct()
  datrdb$FM<-datrdb$FM[rep(1,nrow(pipo)),]%>%
	  mutate(FMid=pipo$newFMid,
		 FMclass=pipo$len,
		 FMnumberAtUnit=pipo$n,
		 FMtype="Length",
		 SAid=pipo$newSAid)
  #SA
  pipo<-tmp%>%transmute(newSAid,newSSid,spp,gear,SAwtot,SAwsamp,SAnsamp,SAntot,SAsel,SAname)%>%distinct()
  datrdb$SA<-datrdb$SA[rep(1,nrow(pipo)),]%>%
	  mutate(SAid=pipo$newSAid,
		 SAsequenceNumber=NA,
		 SAspeciesCode=NA,
		 SAspeciesCodeFAO=pipo$spp,
		 SApresentation=NA,
		 SAspecimensState=NA,
		 SAcatchCategory="all",
		 SAsex="U",
		 SAmetier6=pipo$gear,
		 SAgear=pipo$gear,
		 SAunitType="Haul",
		 SAnumberTotal=pipo$SAntot,
		 SAnumberSampled=pipo$SAnsamp,
		 SAselectionMethod="Census",
		 SAunitName=pipo$SAname,
		 SAtotalWeightMeasured=pipo$SAwtot,
		 SAsampleWeightMeasured=pipo$SAwsamp,
		 SSid=pipo$newSSid)
  #SS
  pipo<-tmp%>%transmute(newSSid,newFOid,SSntot,SSnsamp,SSsel,SSname,SLname=paste0(year,gear))%>%distinct()
  datrdb$SS<-datrdb$SS[rep(1,nrow(pipo)),]%>%
	  mutate(SSid=pipo$newSSid,
		 SSsequenceNumber=NA,
		 SSobservationActivityType="haul",
		 SScatchFraction="All",
		 SSspeciesListName=pipo$SLname,
		 SSnumberTotal=pipo$SSntot,
		 SSnumberSampled=pipo$SSnsamp,
		 SSselectionMethod="Census",
		 SSunitName=pipo$SSname,
		 FOid=pipo$newFOid)
  #SL
  pipo<-tmp%>%transmute(SLname=paste0(year,gear),year,spp)%>%distinct()
  datrdb$SL<-datrdb$SL[rep(1,nrow(pipo)),]%>%
	  mutate(SSid=1:nrow(pipo),
		 SLinstitute="in silico",
		 SLspeciesListName=pipo$SLname,
		 SLyear=pipo$year,
		 SLcatchFraction="All",
		 SLcommercialTaxon=pipo$spp,
		 SLspeciesCode=NA)
  #FO
  pipo<-tmp%>%transmute(newFOid,newFTid,gear,year,FOntot,FOnsamp,FOstratname,FOname,FOsel)%>%distinct()
  datrdb$FO<-datrdb$FO[rep(1,nrow(pipo)),]%>%
	  mutate(FOid=pipo$newFOid,
		 FOsequenceNumber=NA,
		 FOstratification="Y",
		 FOstratumName=pipo$FOstratname,
		 FOvalidity="V",
		 FOendDate=pipo$year,
		 FOmetier6=pipo$gear,
		 FOgear=pipo$gear,
		 FOnumberSampled=pipo$FOnsamp,
		 FOnumberTotal=pipo$FOntot,
		 FOselectionMethod=pipo$FOsel,
		 FOunitName=pipo$FOname,
		 FTid=pipo$newFTid)
  #FT
  pipo<-tmp%>%transmute(newFTid,newVSid,gear,year,FTntot,FTnsamp,FTstratname,FTname,FTsel,VSname,FOntot)%>%distinct()
  datrdb$FT<-datrdb$FT[rep(1,nrow(pipo)),]%>%
	  mutate(FTid=pipo$newFTid,
		 FTsequenceNumber=NA,
		 FTencryptedVesselCode=pipo$VSname,
		 FTstratification="Y",
		 FTstratumName=pipo$FTstratname,
		 FTnumberOfHauls=pipo$FOntot,
		 FTdepartureDate=NA,
		 FTarrivalDate=pipo$year,
		 FTnumberTotal=pipo$FTntot,
		 FTnumberSampled=pipo$FTnsamp,
		 FTselectionMethod=pipo$FTsel,
		 FTunitName=pipo$FTname,
		 VSid=pipo$newVSid)
  #VS
  pipo<-tmp%>%transmute(newVSid,newSDid,VSname,VSstratname,VSntot,VSnsamp,VSsel)%>%distinct()
  datrdb$VS<-datrdb$VS[rep(1,nrow(pipo)),]%>%
	  mutate(VSid=pipo$newVSid,
		 VSsequenceNumber=NA,
		 VSencryptedVesselCode=pipo$VSname,
		 VSstratification="Y",
		 VSstratumName=pipo$VSstratname,
		 VSnumberTotal=pipo$VSntot,
		 VSnumberSampled=pipo$VSnsamp,
		 VSselectionMethod=pipo$VSsel,
		 VSunitName=pipo$VSname,
		 SDid=pipo$newSDid)
  #VD
  pipo<-tmp%>%transmute(VSname,year)%>%distinct()
  datrdb$VD<-datrdb$VD[rep(1,nrow(pipo)),]%>%
	  mutate(VDid=pipo$VSname,
		 VDyear=pipo$year,
		 VDencryptedVesselCode=pipo$VSname,
		 VDlengthCategory=NA)
  #SD and SE
  pipo<-tmp%>%select(newSDid,newDEid,year)%>%distinct()
  datrdb$SD<-datrdb$SD[rep(1,nrow(pipo)),]%>%
	  mutate(SDid=pipo$newSDid,
		 SDinstitution="in silico",
		 DEid=pipo$newDEid)
  datrdb$DE<-datrdb$DE[rep(1,nrow(pipo)),]%>%
	  mutate(DEid=pipo$newDEid,
		 DEsamplingScheme="simulation census",
		 DEsamplingSchemeType="census",
		 DEyear=pipo$year)
  #compute a population data data
  cl<-datsim%>%group_by(year,gear,spp)%>%summarise(w=sum(wspp))%>%ungroup()
  clrdb<-cl%>%transmute(Clid=1:nrow(cl),
	  CLrecType="CL", CLdTypSciWeig="Official",
	  CLdSouSciWeig="Combination of official data",
	  CLsampScheme="census", CLdSouLanVal="Other",
	  CLlanCou="ZZ", CLvesFlagCou="ZZ",
	  CLyear=year, CLquar=NA, CLmonth=NA,
	  CLarea=NA, CLstatRect=NA,CLgsaSubarea=NA,
	  CLjurisdArea=NA, CLeconZoneIndi=NA,
	  CLspecCode=NA, CLspecFAO=cl$spp,
	  CLlandCat="HuC", CLcatchCat="Lan",
	  CLsizeCatScale=NA, CLsizeCat=NA,
	  CLnatFishAct=NA, CLmetier6=cl$gear,
	  CLIBmitiDev=NA, CLloc=NA,
	  CLvesLenCat=NA, CLfishTech=NA,
	  CLdeepSeaReg=NA, CLoffWeight=cl$w,
	  CLsciWeight=NA, CLexpDiff=NA,
	  CLtotOffLanVal=NA, CLnumUniqVes=NA,
	  CLsciLanRSE=NA, CLvalRSE=NA,
	  CLsciLanQualBias=NA)
  return(list(cl=data.frame(clrdb),samp=datrdb))
}#}}}
```

# Outputs

Data are converted and saved.

```{r output}
#pop data rdbes
rezpop<-sim2prepsim(datsim,rep(T,nrow(datsim)))
rezpop<-prepsim2rdbes(rezpop,datrdb,rep(T,nrow(rezpop)))
saveRDS(rezpop$samp,file="../outputs/datrdbsimpop.rds")
saveRDS(rezpop$cl,file="../outputs/datclrdbsim.rds")

#sim data rdbes
rezsamp<-sim2prepsim(datsim,datsim$FOsamp&datsim$SPPsamp)

#rezsamp%>%select(FOsamp,SPPsamp,propsamp,newFOid,newSAid,spp)%>%View()
rezsamp<-prepsim2rdbes(rezsamp,datrdb,datsim$FOsamp&datsim$SPPsamp)
saveRDS(rezsamp$samp,file="../outputs/datrdbsimsamp.rds")

```

`

# R session information

```{r Rsession,cache=TRUE,eval=T,echo=T,cache=T,include=T}
sessioninfo::session_info()
```

# References
